\bf Title here

### 2020-09-10 paper/paper.Rmd
update eixsting alogrithm: start the first algo by search_better; tune the language

Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient ascent.

@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$. The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path. A new round of search continues to find a better basis after the interpolation finishes. The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$. When the search needs to sample more than $c$ number of basis to find an accepted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration. If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops. The algorithm is summarised in Algorithm \ref{random-search} for one iteration. [mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
    \Output{$\mathbf{A}_{l}$}
  initialisation\;
  Set $l = 1$ and $c = 0$\;
  \While{$l < l_{\max}$}{
    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthonormalise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
    \eIf{$I_{l} > I_{\text{cur}}$}{
      \KwRet{$\mathbf{A}_{l}$} \;
      }{
      $c = c + 1$\;
      }
    $l = l + 1$\;
  }
  We repeat this loop and if $c > 30$, half the $\alpha$. Reset $c$ to zero.
  \caption{random search}
  \label{random-search}
\end{algorithm}

Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] modifies `search_better` based on a non-increasing cooling scheme $T(j)$. Given an initial $T_0$, the temperature at iteration $k$ is defined as $T(j) = \frac{T_0}{\log(j + 1)}$. When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis. This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space. This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e. a one dimensional projection of the dataset with two informative variables). The algorithm can be written as replacing line 5-11 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.

\begin{algorithm}
\SetAlgoLined
    Compute $I_{j} = f(\mathbf{A}_{j})$ and $T(k) = \frac{T_0}{\log(j + 1)}$\;
      \eIf{$I_{j} > I_{\text{cur}}$}{
        \KwRet{$\mathbf{A}_{j}$} \;
      }{
        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{j}}{T(j)}\right],1\right\}$\;
        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
        \If{$P > U$}{
           \KwRet{$\mathbf{A}_{j}$} \;
        }
      }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

@cook1995grand used a gradient ascent algorithm with pseudo-derivative. Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the gradient. The $\delta$ parameter is usually set to be tiny to ensure the geodesic direction is a good local approximation to the gradient. Once the searching direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic as the candidate. $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates. Algorithm \ref{search-geodesic} summarises the steps in geodesic search.

\begin{algorithm}
\SetAlgoLined
\SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $j_{\max}$, $n$, $\delta$}
    \Output{$\mathbf{A}_{**}$}
  initialisation\;
  Set $j = 1$\;
  \While{$j < j_{\max}$}{
    Generate $2n$ bases in $n$ random directions: $\mathbf{A}_{j}: \mathbf{A}_{j+9}$ within a small neighbourhood $\delta$\;
    Find the direction with the largest index value: $\mathbf{A}_{*}$ where $j < * < j+ 9$\;
    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
    Find $\mathbf{A}_{**}$ on the geodesic $\mathcal{G}$ that has the largest index value \;
    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
      \If{$p_{\text{diff}} > 0.001$}{
         \KwRet{$\mathbf{A}_{**}$} \;
      }
    $j = j + 1$\;
  }
  \caption{search geodesic}
  \label{search-geodesic}
\end{algorithm}

>>>>

Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate bases in the neighbourhood, whilst `search_geodesic` is an analogue of gradient ascent on the manifold of the projection basis.

`search_better` is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$. $\mathbf{A}_{l}$ is then orthogonalised to ensure the orthonormal constraint is fulfilled. When a basis is found with index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the search terminates and outputs the basis for guided tour to construct an interpolation path. The next iteration of search begins after adjusting $\alpha$ by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$. Another termination condition is when the maximum number of iteration $l_{\max}$ is reached. A slightly different cooling scheme has been proposed by @posse1995projection to include a halving parameter $c$. Rather than reducing the radius of the searching neighbourhood, $\alpha$, at each iteration, Posse's design only adjust $\alpha$ if the last search takes more than $c$ times to find an accepted basis to avoid the searching space being reduced too fast. The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.  [mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
    \Output{$\mathbf{A}_{l}$}
  initialisation\;
  Set $l = 1$\;
  \While{$l < l_{\max}$}{
    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
    \If{$I_{l} > I_{\text{cur}}$}{
      \KwRet{$\mathbf{A}_{l}$} \;
      }
    $l = l + 1$\;
  }
  \caption{random search}
  \label{random-search}
\end{algorithm}

Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme $T(l)$. Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$. When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis. This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space. This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e. a one dimensional projection of the dataset with two informative variables). The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.

\begin{algorithm}
\SetAlgoLined
    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
      \eIf{$I_{l} > I_{\text{cur}}$}{
        \KwRet{$\mathbf{A}_{l}$} \;
      }{
        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
        \If{$P > U$}{
           \KwRet{$\mathbf{A}_{l}$} \;
        }
      }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

@cook1995grand used a gradient ascent algorithm on the manifold of the projection bases. In gradient ascent, one first find the direction for improvment via computing the gradient information. In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$. A geodesic is then constructed using the current basis and one of the $2n$ bases that have the highest index value. If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction. In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction. In `search_geodesic`, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from $-\pi/4$ to $\pi/4$. The optima $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value. As above, another termination condition is when $l_{\max}$ is reached.  Algorithm \ref{search-geodesic} summarises the steps in geodesic search.

\begin{algorithm}
\SetAlgoLined
\SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n = 5$, $\delta$}
    \Output{$\mathbf{A}_{**}$}
  initialisation\;
  Set $l = 1$\;
  \While{$l < l_{\max}$}{
    Generate $2n$ bases in a small neighbourhood, $\delta$, of $\mathbf{A}_{\text{cur}}$ and ensure orthogonality \;
    Find the one with the largest index value: $\mathbf{A}_{*}$\;
    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
    Optimise the index value on the geodesic $\mathcal{G}$ over a 90 degree window to produce the optima $\mathbf{A}_{**}$  \;
    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
      \If{$p_{\text{diff}} > 0.001$}{
         \KwRet{$\mathbf{A}_{**}$} \;
      }
    $l = l + 1$\;
  }
  \caption{search geodesic}
  \label{search-geodesic}
\end{algorithm}

### 2020-09-10 paper/paper.tex
update eixsting alogrithm: start the first algo by search_better; tune the language

Below we introduce three possible algorithms: \texttt{search\_better}
and \texttt{search\_better\_random} are derivative free methods that
sample candidate basis in the neighbourhood whilst
\texttt{search\_geodesic} approximates the gradient information and acts
as a non-derivative version of the gradient ascent.

\citet{posse1995projection} proposed a random search algorithm that
samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the
current basis \(\mathbf{A}_{\text{cur}}\) by
\(\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\),
where \(\alpha\) controls the radius of the sampling neighbourhood and
\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
same dimension as \(\mathbf{A}_{\text{cur}}\). The optimiser keeps
sampling bases near the current basis until it finds one with higher
index value than the current basis and then outputs it for guided tour
to construct the interpolation path. A new round of search continues to
find a better basis after the interpolation finishes. The halving
parameter \(c\) with default value of 30 is designed to adjust the
searching neighbourhood \(\alpha\). When the search needs to sample more
than \(c\) number of basis to find an accepted basis, the neighbourhood
parameter \(\alpha\) will be reduced by half in the next iteration. If
the optimiser can't find a better basis within the maximum number of
tries \(l_{\max}\), the algorithm stops. The algorithm is summarised in
Algorithm \ref{random-search} for one iteration. {[}mention
orthonormalise to ensure the constraint is fulfilled; don't use
derivative information but a random search{]}

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
    \Output{$\mathbf{A}_{l}$}
  initialisation\;
  Set $l = 1$ and $c = 0$\;
  \While{$l < l_{\max}$}{
    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthonormalise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
    \eIf{$I_{l} > I_{\text{cur}}$}{
      \KwRet{$\mathbf{A}_{l}$} \;
      }{
      $c = c + 1$\;
      }
    $l = l + 1$\;
  }
  We repeat this loop and if $c > 30$, half the $\alpha$. Reset $c$ to zero.
  \caption{random search}
  \label{random-search}
\end{algorithm}

Simulated annealing
\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization}
modifies \texttt{search\_better} based on a non-increasing cooling
scheme \(T(j)\). Given an initial \(T_0\), the temperature at iteration
\(k\) is defined as \(T(j) = \frac{T_0}{\log(j + 1)}\). When a candidate
basis fails to have an index value larger than the current basis,
simulated annealing gives it a second chance to be accepted with
probability
\[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}\]
where \(I_{(\cdot)}\) denotes the index value of a given basis. This
implementation allows the algorithm to jump out of a local maximum and
enables a more holistic search of the whole parameter space. This
feature is particularly useful when the dimension of the projected space
is smaller than the number of informative variables in the dataset
(i.e.~a one dimensional projection of the dataset with two informative
variables). The algorithm can be written as replacing line 5-11 of
Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.

\begin{algorithm}
\SetAlgoLined
    Compute $I_{j} = f(\mathbf{A}_{j})$ and $T(k) = \frac{T_0}{\log(j + 1)}$\;
      \eIf{$I_{j} > I_{\text{cur}}$}{
        \KwRet{$\mathbf{A}_{j}$} \;
      }{
        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{j}}{T(j)}\right],1\right\}$\;
        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
        \If{$P > U$}{
           \KwRet{$\mathbf{A}_{j}$} \;
        }
      }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

\citet{cook1995grand} used a gradient ascent algorithm with
pseudo-derivative. Instead of computing the actual gradient of the index
function with respect to the projection basis matrix,
\texttt{search\_geodesic} samples \(2n\) bases that are randomly
generated on a uniform open ball with the radius controlled by the
\(\delta\) parameter and finds the most promising one to construct the
geodesic direction as an approximation to the gradient. The \(\delta\)
parameter is usually set to be tiny to ensure the geodesic direction is
a good local approximation to the gradient. Once the searching direction
is determined, the optimiser finds the best projection
\(\mathbf{A}_{**}\) on the geodesic as the candidate.
\(\mathbf{A}_{**}\) is outputted for the current iteration if the
percentage change in the index value between \(\mathbf{A}_{**}\) and
\(\mathbf{A}_{\text{cur}}\) is greater than a threshold value or the
algorithm repeats the above steps until \(l_{\max}\) is reached and the
searching terminates. Algorithm \ref{search-geodesic} summarises the
steps in geodesic search.

\begin{algorithm}
\SetAlgoLined
\SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $j_{\max}$, $n$, $\delta$}
    \Output{$\mathbf{A}_{**}$}
  initialisation\;
  Set $j = 1$\;
  \While{$j < j_{\max}$}{
    Generate $2n$ bases in $n$ random directions: $\mathbf{A}_{j}: \mathbf{A}_{j+9}$ within a small neighbourhood $\delta$\;
    Find the direction with the largest index value: $\mathbf{A}_{*}$ where $j < * < j+ 9$\;
    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
    Find $\mathbf{A}_{**}$ on the geodesic $\mathcal{G}$ that has the largest index value \;
    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
      \If{$p_{\text{diff}} > 0.001$}{
         \KwRet{$\mathbf{A}_{**}$} \;
      }
    $j = j + 1$\;
  }
  \caption{search geodesic}
  \label{search-geodesic}
\end{algorithm}

>>>>

Below we introduce three possible algorithms: \texttt{search\_better}
and \texttt{search\_better\_random} are derivative free methods that
sample candidate bases in the neighbourhood, whilst
\texttt{search\_geodesic} is an analogue of gradient ascent on the
manifold of the projection basis.

\texttt{search\_better} is a random search device that samples a
candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
basis \(\mathbf{A}_{\text{cur}}\) by
\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
where \(\alpha\) controls the radius of the sampling neighbourhood and
\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
then orthogonalised to ensure the orthonormal constraint is fulfilled.
When a basis is found with index value higher than the current basis
\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
for guided tour to construct an interpolation path. The next iteration
of search begins after adjusting \(\alpha\) by a cooling parameter:
\(\alpha_{j+1} = \alpha_j * \text{cooling}\). Another termination
condition is when the maximum number of iteration \(l_{\max}\) is
reached. A slightly different cooling scheme has been proposed by
\citet{posse1995projection} to include a halving parameter \(c\). Rather
than reducing the radius of the searching neighbourhood, \(\alpha\), at
each iteration, Posse's design only adjust \(\alpha\) if the last search
takes more than \(c\) times to find an accepted basis to avoid the
searching space being reduced too fast. The algorithm of
\texttt{search\_better} is summarised in Algorithm \ref{random-search}.
{[}mention orthonormalise to ensure the constraint is fulfilled; don't
use derivative information but a random search{]}

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
    \Output{$\mathbf{A}_{l}$}
  initialisation\;
  Set $l = 1$\;
  \While{$l < l_{\max}$}{
    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
    \If{$I_{l} > I_{\text{cur}}$}{
      \KwRet{$\mathbf{A}_{l}$} \;
      }
    $l = l + 1$\;
  }
  \caption{random search}
  \label{random-search}
\end{algorithm}

Simulated annealing
\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses
the same sampling process as \texttt{search\_better} but allow a
probabilistic acceptance of a basis with lower index value based on a
cooling scheme \(T(l)\). Given an initial \(T_0\), the temperature at
iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
candidate basis fails to have an index value larger than the current
basis, simulated annealing gives it a second chance to be accepted with
probability
\[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\]
where \(I_{(\cdot)}\) denotes the index value of a given basis. This
implementation allows the algorithm to jump out of a local maximum and
enables a more holistic search of the whole parameter space. This
feature is particularly useful when the dimension of the projected space
is smaller than the number of informative variables in the dataset
(i.e.~a one dimensional projection of the dataset with two informative
variables). The algorithm can be written as replacing line 5-8 of
Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.

\begin{algorithm}
\SetAlgoLined
    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
      \eIf{$I_{l} > I_{\text{cur}}$}{
        \KwRet{$\mathbf{A}_{l}$} \;
      }{
        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
        \If{$P > U$}{
           \KwRet{$\mathbf{A}_{l}$} \;
        }
      }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

\citet{cook1995grand} used a gradient ascent algorithm on the manifold
of the projection bases. In gradient ascent, one first find the
direction for improvment via computing the gradient information. In
\texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny
neighbourhood of the current basis, controlled by the neighbourhood
parameter \(\delta\). A geodesic is then constructed using the current
basis and one of the \(2n\) bases that have the highest index value. If
the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed
is an analogue of the gradient information in the curved space and can
work as the searching direction. In gradient ascent, the next step is to
conduct a line search to find the best improvement along the gradient
direction. In \texttt{search\_geodesic}, this step is replaced by
optimising the index value along the geodesic direction constructed
before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\). The optima
\(\mathbf{A}_{**}\) is outputted for the current iteration if the
percentage change in the index value between \(\mathbf{A}_{**}\) and
\(\mathbf{A}_{\text{cur}}\) is greater than a threshold value. As above,
another termination condition is when \(l_{\max}\) is reached. Algorithm
\ref{search-geodesic} summarises the steps in geodesic search.

\begin{algorithm}
\SetAlgoLined
\SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n = 5$, $\delta$}
    \Output{$\mathbf{A}_{**}$}
  initialisation\;
  Set $l = 1$\;
  \While{$l < l_{\max}$}{
    Generate $2n$ bases in a small neighbourhood, $\delta$, of $\mathbf{A}_{\text{cur}}$ and ensure orthogonality \;
    Find the one with the largest index value: $\mathbf{A}_{*}$\;
    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
    Optimise the index value on the geodesic $\mathcal{G}$ over a 90 degree window to produce the optima $\mathbf{A}_{**}$  \;
    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
      \If{$p_{\text{diff}} > 0.001$}{
         \KwRet{$\mathbf{A}_{**}$} \;
      }
    $l = l + 1$\;
  }
  \caption{search geodesic}
  \label{search-geodesic}
\end{algorithm}
